Methods paper for erosion

### 2017-11-08 docs/methods/erosion_methods.tex
Working on theta-L section.

@@

We now introduce the tangent angle, $\theta$, defined through the relation
\begin{equation}
\left( \pderiv{x}{s}, \pderiv{y}{s} \right) = \left(\cos \theta, \sin \theta \right)
\end{equation}
where $s$ is arc length. We also introduce the normalized arc length $\alpha = s / L$. In terms of these new variables, the curvature can be expressed as
\begin{equation}
\kappa = \frac{1}{L} \pderiv{\theta}{\alpha}
\end{equation}
Substitution into Eq.~(\ref{eqn:Vn}) gives
\begin{equation}
\label{eqn:Vn2}
\Vn = \atau + \frac{\eps}{L \log \left(2\pi/L \right)}  \left(\pderiv{\theta}{\alpha} - 2 \pi \right)
\end{equation}
The main idea is to introduce an artificial tangential velocity that keeps collocation points equally spaced with respect to arc length; i.e.~the grid of $\alpha$ should always remain equally spaced. The tangential velocity that does the trick is determined by \cite{hou-low-she1994}
\begin{equation}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}
\end{equation}
The interface evolution is then given by
\begin{equation}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{equation}
Upon transforming to the variables $\theta(\alpha,t)$ and $L(t)$, we obtain the evolution equations
\begin{align}
& \tderiv{L}{t} = - \mean{\thalpha \Vn} \\
& \pderiv{\theta}{t} = \frac{1}{L} \left( \pderiv{\Vn}{\alpha} + \thalpha \Vs \right)
\end{align}
Inserting Eq.~(\ref{eqn:Vn2}) for the normal velocity gives
\begin{equation}
\pderiv{\theta}{t} = \frac{\eps}{L^{2} \log \left(2\pi/L \right)} \ppd{\theta}{\alpha} 
+ \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right)
\end{equation}
For simplicity, we introduce the variables
\begin{align}
& M = - \mean{\thalpha \Vn} \\
& N = \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
& \elfun = \frac{1}{L^{2} \log \left(2\pi/L \right)}
\end{align}
Then the evolution equations can be expressed as
\begin{align}
& \tderiv{L}{t} = M \\
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + N
\end{align}
We treat $\theta$ spectrally
\begin{equation}
\theta(\alpha,t) = \sum_{k = -\infty}^{\infty} \thhat_k e^{2\pi i k \alpha}
\end{equation}
Then we get
\begin{equation}
\tderiv{\thhat_k}{t} +  4 \eps \pi^2 k^2  \elfun \thhat_k = \hat{N}_k
\end{equation}
We introduce the integrating factor
\begin{equation}
\mu = \exp \left( 4 \eps \pi^2 k^2 \int \elfun dt \right)
\end{equation}
so that the solution can be represented as
\begin{align}
\thhat_k(t) & = \frac{1}{\mu} \int^t \mu \hat{N}_k dt'
= \thhat_k(0) + \int_0^t \exp \left( -4 \eps \pi^2 k^2 \int_{t'}^t \elfun dt'' \right) \hat{N}_k \, dt'
% LONGER FORM
%& = \exp \left( -4 \eps \pi^2 k^2 \int_0^t \elfun(L) ds \right) \, 
%\int_0^t \exp \left( 4 \eps \pi^2 k^2 \int_0^s \elfun(L) ds' \right) \hat{N}_k ds \\
\end{align}
where $t'$ and $t''$ are simply dummy variables for integration. We can get the exact relationship
\begin{equation}
\thhat_k(t + \Delta t) = \thhat_k(t) 
+ \int_t^{t+\Delta t} \exp \left( -4 \eps \pi^2 k^2 \int_{t'}^{t+\Delta t} \elfun dt'' \right) \hat{N}_k \, dt'
\end{equation}

>>>>

\todo[inline]{Bryan, I think you should be first author on this paper.}

@@

We now introduce the tangent angle, $\theta$, defined through the relation
\begin{equation}
\left( \pderiv{x}{s}, \pderiv{y}{s} \right) = \left(\cos \theta, \sin \theta \right)
\end{equation}
where $s$ is arc length. We also introduce the normalized arc length $\alpha = s / L$. In terms of these new variables, the curvature can be expressed as
\begin{equation}
\kappa = \frac{1}{L} \pderiv{\theta}{\alpha}
\end{equation}
Substitution into Eq.~(\ref{eqn:Vn}) gives
\begin{equation}
\label{eqn:Vn2}
\Vn = \atau + \frac{\eps}{L \log \left(2\pi/L \right)}  \left(\pderiv{\theta}{\alpha} - 2 \pi \right)
\end{equation}

The main idea is to introduce an artificial tangential velocity that keeps collocation points equally spaced with respect to arc length; i.e.~the grid of $\alpha$ should always remain equally spaced. The tangential velocity that does the trick is determined by \cite{hou-low-she1994}
\begin{equation}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}
\end{equation}
The interface evolution is then given by
\begin{equation}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{equation}
Upon transforming to the variables $\theta(\alpha,t)$ and $L(t)$, we obtain the evolution equations
\begin{align}
\label{eqn:Levo0}
& \tderiv{L}{t} = - \mean{\thalpha \Vn} \\
\label{eqn:thetaevo0}
& \pderiv{\theta}{t} = \frac{1}{L} \left( \pderiv{\Vn}{\alpha} + \thalpha \Vs \right)
\end{align}
Inserting Eq.~(\ref{eqn:Vn2}) for the normal velocity gives
\begin{equation}
\pderiv{\theta}{t} = \frac{\eps}{L^{2} \log \left(2\pi/L \right)} \ppd{\theta}{\alpha} 
+ \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right)
\end{equation}

For simplicity, we introduce the variables
\begin{align}
\label{eq:M}
& M = - \mean{\thalpha \Vn} \\
\label{eq:N}
& N = \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
& \elfun = \frac{1}{L^{2} \log \left(2\pi/L \right)}
\end{align} 
Then Eq.~(\ref{eqn:thetaevo0}) can be rewritten as
\begin{equation}
\label{eqn:thetaevo1}
\pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + N
\end{equation}
Note that the $\elfun$ depends only on $L(t)$, and so the highest order derivative in Eq.~(\ref{eqn:thetaevo1}) is {\em linear} in the variable $\alpha$. This is the key to the {\thL} method, and allows us to use very stable, implicit methods for the stiff term. We will evolve $\theta$ in spectral space, and so introduce the transforms
\begin{align}
& \theta(\alpha,t) = \sum_{k = -\infty}^{\infty} \thhat_k e^{2\pi i k \alpha} \\
& N(\alpha,t)  = \sum_{k = -\infty}^{\infty} \hat{N}_k e^{2\pi i k \alpha}
\end{align}
Equation (\ref{eqn:thetaevo1}) gives a set of ODEs for the coefficients
\begin{equation}
\tderiv{\thhat_k}{t} +  4 \eps \pi^2 k^2  \elfun \thhat_k = \hat{N}_k
\end{equation}
We introduce the integrating factor
\begin{equation}
\label{eq:mu}
\mu_k = \exp \left( 4 \eps \pi^2 k^2 \int \elfun \, dt \right)
\end{equation}
Then the evolution equations for $L$ and $\theta$ can be recast as the coupled system of ODEs
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = M \\
\label{eqn:thetaevo2}
& \tderiv{}{t}\left( \mu_k \theta_k \right) = \mu_k \hat{N}_k
\end{align}
We solve system (\ref{eqn:Levo2})--(\ref{eqn:thetaevo2}) for the variables $L$ and $\{ \mu_k \theta_k \}$ with a 4th-order Runge-Kutta method. At each stage, we must recover $\theta(\alpha,t)$ from  $\{ \mu_k \theta_k \}$ in order to compute the shear stress as needed in Eqs.~(\ref{eq:M}) and (\ref{eq:N}) for $M$ and $N$. For this, we compute $\mu_k$ from Eq.~(\ref{eq:mu}) using the trapezoid rule, then recover $\theta(\alpha, t)$ via a Gaussian filter....
